---
title: "Work with data types"
format: html
---

# Instructions

This week, you worked with three new data types in R -- factors, strings, and date-times -- using the forcats, stringr, and lubridate packages, respectively.

In this milestone, you'll work with the treatment type data contained in `mdrd` to investigate whether diet type affects change in GFR over time. 

To learn more about the variables contained in the data set(s), refer to the data dictionary: https://rsacdn.link/milestones/internal/v2/tidyverse/renal-treatment/assets/mdrd_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false
```

## Data types

In this milestone, you'll use stringr and forcats to wrangle the treatment type data -- separating and recoding it into tidy `diet` and blood pressure (`bp`) variables for the very low-protein and low-protein diet groups. 

## Recreation

### Part 1 - Import

To begin, read the data from `mdrd.csv` which lives in your `data/` directory to an object named `mdrd`:

```{r}
#| label: recreation-import

```

You will be working with a tidied version of the mdrd data set. Run the code chunk below, which contains code from milestone 5, to create `mdrd_tidy`. 

```{r}
#| label: recreation-create-tidied
mdrd_tidy <- 
  mdrd %>%
  pivot_longer(
    cols = starts_with("diet"),
    names_to = "treatment",
    values_to = "value"
  ) %>%
  filter(value == 1) %>%
  select(-value)

mdrd_tidy
```

### Part 2 - Working with data types

Run the code below to see a table.

```{r}
#| label: recreate-this-1
#| message: false
solution <- readr::read_csv(
  "data/milestone05.csv",
  col_types = cols(
    diet = col_factor(c("very low-protein", "low-protein")),
    bp   = col_factor(c("low", "normal"))
  )
)

solution
```

Your task is to use what you've learned about the stringr package to transform `mdrd_tidy` into this table. You will need to:

1. Create two new columns, `diet` and `bp`, by separating the character strings in `treatment`. (*Hint* Look up the help page for `tidyr::separate_wider_delim()`)

2. Remove `"diet"` from the values in `diet`. Also remove `"bp"` from the values in `bp`. (*Hint* Look up the help page for `stringr::str_remove()`)

3. Recode the values in `diet` to make them more informative. Recode the values such that:

    + `"k"` = `"very low-protein"`
    + `"l"` = `"low-protein"`

4. Recode the values in `bp` such that `"norm"` becomes `"normal"`.

(*Hint* Look up the help page for `forcats::fct_recode()`)

Work in the code chunk below. Save the result as `treatments`.

```{r}
#| label: recreation-seasons

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(treatments, solution, tolerance = 1e-4, ignore_attr = "class")
```

## Extension

Using the code chunk below, investigate a research question about this data, using the skills you have learned this week and in previous weeks of Academy. Some ideas:

1. Is there an interaction effect between diet and blood pressure group?
2. Does diet or blood pressure group affect whether or not a patient dropped out of the study?
3. [any other research question of interest]


Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. If your data contains ...
  ... date-time data, use a lubridate function to correctly format your date-time variables
  ... string data, use a stringr function to manipulate your string variables
  ... categorical data, use a forcats function to modify the factor levels of your categorical variables (reorder, recode, etc.)
3. Using a table or graph, present a meaningful subset of your modified data


```{r}
#| label: extension

```
