---
title: "Work with data types"
format: html
---

# Instructions

This week, you worked with three new data types in R -- factors, strings, and date-times -- using the forcats, stringr, and lubridate packages, respectively.

In this milestone, you'll work with the treatment type data contained in `mdrd` to investigate whether diet type affects change in GFR over time. 

To learn more about the variables contained in the data set(s), refer to the data dictionary: https://rsacdn.link/milestones/internal/v2/tidyverse/renal-treatment/assets/mdrd_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false
```

## Data types

In this milestone, you'll use stringr and forcats to investigate whether change in GFR over time varies between the very low-protein diet and the low-protein diet. 

## Recreation

### Part 1 - Import

To begin, read the data from `mdrd.csv` which lives in your `data/` directory to an object named `mdrd`:

```{r}
#| label: recreation-import

```

You will be working with a tidied version of the mdrd data set. Run the code chunk below, which contains code from milestone 5, to create `mdrd_tidy`. 

```{r}
#| label: recreation-create-tidied
mdrd_tidy <- 
  mdrd %>%
  pivot_longer(
    cols = starts_with("diet"),
    names_to = "treatment",
    values_to = "value"
  ) %>%
  filter(value == 1) %>%
  select(-value)

mdrd_tidy
```

### Part 2 - Working with data types

Run the code below to see a table.

```{r}
#| label: recreate-this-1
#| message: false
solution <- readr::read_csv(
  "data/milestone05.csv",
  col_types = cols(
    diet = col_factor(c("very low-protein", "low-protein")),
    bp   = col_factor(c("low", "normal"))
  )
)

solution
```

Your task is to use what you've learned about the stringr package to transform `mdrd_tidy` into this table. You will need to:

1. Create two new columns, `diet` and `bp`, by separating the character strings in `treatment`. (*Hint* Look up the help page for `tidyr::separate_wider_delim()`)

2. Remove `"diet"` from the values in `diet`. Also remove `"bp"` from the values in `bp`. (*Hint* Look up the help page for `stringr::str_remove()`)

3. Recode the values in `diet` to make them more informative. Recode the values such that:

    + `"k"` = `"very low-protein"`
    + `"l"` = `"low-protein"`

4. Recode the values in `bp` such that `"norm"` becomes `"normal"`.

(*Hint* Look up the help page for `forcats::fct_recode()`)

Work in the code chunk below. Save the result as `treatments`; we will use it in Part 3.

```{r}
#| label: recreation-seasons

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(treatments, solution, tolerance = 1e-4, ignore_attr = "class")
```

### Part 3 - The role of factor orders in visualizations

Now that you have `diet` and `bp` variables, you can compute and visualize summaries for each diet and blood pressure group in the same manner as previous milestones. Run the chunk below to view a summary table.

```{r}
#| label: summary
treatments_summary <- 
  treatments %>%
  group_by(diet, bp, ptid) %>%
  summarize(
    gfr_slope = (last(gfr, order_by = months) - first(gfr, order_by = months)) / max(months)
  ) %>%
  filter(!is.nan(gfr_slope)) %>% 
  ungroup()

treatments_summary
```

After you have run the chunk above to create `treatments_summary`, run the chunk below to visualize the data of `treatments_summary`.

```{r}
#| label: summary-plot
#| eval: false
ggplot(treatments_summary, aes(x = diet, y = gfr_slope)) +
  geom_boxplot()
```

This is a nice plot, but the order of `diet` on the x axis is arbitrary (alphabetical). We may want to give the `diet` a more principled order before we publish this plot.

#### Visualization

Using the File Explorer, navigate to the `images/` directory, then click on `milestone05.png` to open an image.

Notice that "very low-protein" now appears to the left of "low-protein" on the x-axis. Your task is to recreate this plot.

Use the chunk below to recreate this plot. *Hint: turn the `diet` variable into a factor with a set order.*

```{r}
#| label: recreation-visualize-1

```

## Extension

Using the code chunk below, investigate a research question about this data, using the skills you have learned this week and in previous weeks of Academy. Some ideas:

1. Is there an interaction effect between diet and blood pressure group?
2. Does diet or blood pressure group affect whether or not a patient dropped out of the study?
3. [any other research question of interest]


Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. If your data contains ...
  ... date-time data, use a lubridate function to correctly format your date-time variables
  ... string data, use a stringr function to manipulate your string variables
  ... categorical data, use a forcats function to modify the factor levels of your categorical variables (reorder, recode, etc.)
3. Using a table or graph, present a meaningful subset of your modified data


```{r}
#| label: extension

```
